---
title: "tfm Carme Mora"
author: "Carme Mora"
date: "9/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library("ggthemes")
library(readxl)

sourcedata <- read.table("dades/dades_coloms.csv", sep=";",stringsAsFactors=F, header = TRUE, dec = ".")

sourcedata <- sourcedata[sourcedata$ecoAll %in% c("TFI", "TFC", "AFC", "AFI", "TOC","GFC", "GFI"),]

```
# **gràfics inicials** #
## Question 1:	The increase in brain size is a followed patern by insular columbiforms as predicted by Sayol et al. (2018)?

Scatter plot mida absoluta
```{r}
ggplot(sourcedata, aes(x=log(body.g), y=log(brain.g), color=island)) +
  geom_point(size=1)+ 
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Log body size (g)",y = "Log brain size (g)") +
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))
```

Box plot mida relativa

```{r}
sourcedata$brain.res <- lm(log(brain.g) ~ log(body.g), data=sourcedata)$resid

ggplot(sourcedata, aes(x= island, y=brain.res, color= island)) +
  geom_boxplot() +
  labs(x = "Continent/island",y = "Brain residuals") +
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))
```

## Question 2: Do arboreal foraging pigeons have larger brains than terrestrial foraging ones?
scatter plot

```{r}
ggplot(sourcedata, aes(x=log(body.g), y=log(brain.g), color=foraging)) +
  geom_point(size=1)+ 
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Log body size (g)",y = "Log brain size (g)") +
   theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16)) 
```

relative brain size
```{r}

ggplot(sourcedata, aes(x= foraging, y=brain.res, color= foraging)) +
  geom_boxplot() +
  labs(x = "Foraging strategy",y = "Brain residuals") +
 theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))
```


## Question 3: The observed increase in brain size in the island dwellers is due to the change of feeding pattern? Is the increase in brain size due to insularity and the possible increase of brain size due to the change from terrestrial to arboreal feeding independent or are they additive phenomena?

scatter plot absolute brain size
```{r}

dades5eco <- sourcedata[sourcedata$ecoAll %in% c("TFI", "TFC", "AFC", "AFI", "TOC"),]

ggplot(dades5eco, aes(x=log(body.g), y=log(brain.g), color=ecoAll)) +
  geom_point(size=1)+ 
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Log body size (g)",y = "Log brain size (g)") + 
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))
      
```

boxplot relative brain size

```{r}

ggplot(dades5eco, aes(x= ecoAll, y=brain.res))+
  geom_boxplot(notch=TRUE, notchwidth = 0.5,  color="black", alpha = 0.7, fill = c("red","olivedrab3","darkgreen","blue","purple"), colour = c("red","olivedrab3","darkgreen","blue","purple"))+
  labs(x = "Ecology",y = "Brain residuals") +
   guides(fill=guide_legend(title="Ecological categories")) +
 theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))

```


```{r}
ggplot(sourcedata, aes(x= ecoAll, y=brain.res, color= ecoAll)) +
  geom_boxplot() +
  labs(x = "Ecology",y = "Brain residuals") +
 theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))
```

body size:
only 5 niche categories

```{r}
ggplot(dades5eco, aes(x= ecoAll, y= log(body.g))) +
  geom_boxplot(notch=TRUE, notchwidth = 0.5,  color="black", alpha = 0.7, fill = c("red","olivedrab3","darkgreen","blue","purple"), colour = c("red","olivedrab3","darkgreen","blue","purple"))+
  labs(x = "Ecology",y = "log body mass (g) ") +
  guides(fill=guide_legend(title="Ecological categories")) +
 theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))
```

all niche categories

```{r}
ggplot(sourcedata, aes(x= ecoAll, y= log(body.g), color= ecoAll)) +
  geom_boxplot() +
  labs(x = "Ecology",y = "log body mass") +
 theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 16))
```


```{r,warning=FALSE,message=FALSE}
library("caper")
library("ape")
```

```{r}
phy147 <- read.tree("dades/Columbiformes_jetz_tree_147sps.tre")

sourcedata$ecoAll <- factor(sourcedata$ecoAll, levels = c("TFI", "TFC", "AFC", "AFI", "TOC"))

cdat <- comparative.data(data=sourcedata, phy=phy147, names.col="species.jetz", vcv=TRUE)

sourcedata$brain.res <- lm(log(brain.g) ~ log(body.g), data=sourcedata)$resid

```

## Question 1:	The increase in brain size is a followed patern by insular columbiforms as predicted by Sayol et al. (2018)?  

exploring body size

```{r}
mod1 <- pgls(log(brain.g) ~ log(body.g) + island, cdat, lambda='ML')
summary(mod1)
```

## Question 2: Do arboreal foraging pigeons have larger brains than terrestrial foraging ones?

explorant mida absoluta i factor ecologic
```{r}
mod2<- pgls(log(brain.g) ~ log(body.g) + foraging, cdat, lambda='ML')
summary(mod2)
```
explorar la interacció
```{r}
mod3 <- pgls(log(brain.g) ~ log(body.g) + foraging*island, cdat, lambda='ML')
summary(mod3)
```

## Question 3: The observed increase in brain size in the island dwellers is due to the change of feeding pattern? Is the increase in brain size due to insularity and the possible increase of brain size due to the change from terrestrial to arboreal feeding independent or are they additive phenomena?

exploring body size and ecology

```{r}
mod4 <- pgls(log(brain.g) ~ log(body.g) + ecoAll, cdat, lambda='ML')
summary(mod4)

```

# explorant l'efecte de la mida: 
és significativa en algun model si la fiquem com a variable resposta?

```{r}

mod5 <- pgls(log(body.g) ~ island, cdat, lambda='ML')
summary(mod5)

mod6<- pgls(log(body.g) ~  foraging, cdat, lambda='ML')
summary(mod6)

mod7 <- pgls(log(body.g) ~ foraging*island, cdat, lambda='ML')
summary(mod7)

mod8 <- pgls(log(body.g) ~  ecoAll, cdat, lambda='ML')
summary(mod8)

```
taking AFI as reference
```{r}
sourcedata$ecoAll <- factor(sourcedata$ecoAll, levels = c("AFI", "TFC", "AFC", "TFI", "TOC"))

cdat2 <- comparative.data(data=sourcedata, phy=phy147, names.col="species.jetz", vcv=TRUE)

mod9 <- pgls(log(brain.g) ~ log(body.g) + ecoAll, cdat2, lambda='ML')
summary(mod9)

```

Recontruccions filogenètiques

```{r}
library(phytools)
library(OUwie)

# Fem servir nomes les 5 categories
phy147 <- read.tree("dades/Columbiformes_jetz_tree_147sps.tre")
sourcedata <- read.table("dades/dades_coloms.csv", sep=";",stringsAsFactors=F, header = T)
# Brain residual
sourcedata$brain.res <- lm(log(brain.g) ~ log(body.g), data=sourcedata)$resid
```


```{r}
# Fem servir nomes les 5 categories
cdat <- sourcedata[sourcedata$ecoAll %in% c("AFC","AFI","TFC","TFI","TOC"),]
# Eliminem de l'arbre les especies qu eno fem servir (GFI, TOI, etc)
ctree <- drop.tip(phy147,tip = phy147$tip.label[!phy147$tip.label %in% cdat$species])
```


```{r}
# Stocastic character mapping
ecoAll <- as.character(cdat$ecoAll)
names(ecoAll) <- cdat$species # add the names
length(ecoAll)

fitER <- fitMk(ctree, ecoAll,model="ER",Q="empirical")
fitSYM <- fitMk(ctree, ecoAll,model="SYM",Q="empirical")
fitARD <- fitMk(ctree, ecoAll,model="ARD",Q="empirical")
AIC <- setNames(sapply(list(fitER,fitSYM,fitARD),AIC),c("ER","SYM","ARD"))
AIC

simSYM <- make.simmap(ctree, ecoAll, nsim=10, model = "SYM", Q="empirical") # nsim= number of recontructions 
describe.simmap(simSYM) # Average number of transitions
cols <- c("red","olivedrab3","darkgreen","blue","purple")
names(cols) <- c("AFC","AFI","TFC","TFI","TOC")

# Plot a single simmap
plotSimmap(simSYM[[1]],fsize=0.25,col=cols) 

# Plot all results (piecharts at nodes)
simmaps.obj <- describe.simmap(simSYM,plot=FALSE)
apply(simmaps.obj$count[,2:6],2,function(x){quantile(x,c(0.025,0.975))})
plot(simmaps.obj,fsize=0.35, cex=c(0.5,0.3), nodes="intermediate", ftype="bi", colors=cols)

## Models OUwie
OUdat.brain <- cdat[,c("species","ecoAll","brain.res")] # Species,Category,Phenotype
head(OUdat.brain)

x <- 1 ## Simmap reconstruction number

# Differents models evolutius (BM!,BMS,OU1,OUM)
bm1 <- OUwie(simSYM[[x]], OUdat.brain, model=c("BM1"),simmap.tree=T, mserr="none", algorithm=c("invert"))
bms <- OUwie(simSYM[[x]], OUdat.brain, model=c("BMS"),simmap.tree=T, mserr="none", algorithm=c("invert"))
ou1 <- OUwie(simSYM[[x]], OUdat.brain, model=c("OU1"),simmap.tree=T, mserr="none", algorithm=c("invert"))
oum <- OUwie(simSYM[[x]], OUdat.brain, model=c("OUM"),simmap.tree=T, mserr="none", algorithm=c("invert"))
oumv <- OUwie(simSYM[[x]], OUdat.brain, model=c("OUMV"),simmap.tree=T, mserr="none", algorithm=c("invert"))
ouma <- OUwie(simSYM[[x]], OUdat.brain, model=c("OUMA"),simmap.tree=T, mserr="none", algorithm=c("invert"))

bm1$AIC
bms$AIC
ou1$AIC
oum$AIC
oumv$AIC
ouma$AIC

## Check solutions

# Rate of evolution (Sigma2)
bm1$solution
bms$solution

# Optimum (theta)
oum$theta
oumv$theta

# Strength of selection (Alpha)
ouma$solution

oum$solution
```

```{r}
oumR1 <- OUwie(simSYM[[1]], OUdat.brain, model=c("OUM"),simmap.tree=T, mserr="none", algorithm=c("invert"))
oumR2 <- OUwie(simSYM[[2]], OUdat.brain, model=c("OUM"),simmap.tree=T, mserr="none", algorithm=c("invert"))
oumR3 <- OUwie(simSYM[[3]], OUdat.brain, model=c("OUM"),simmap.tree=T, mserr="none", algorithm=c("invert"))
```

```{r}
oumR1$theta
rec1 <- oumR1$theta[,1] # primera columna, que són els optims.
rec2 <- oumR2$theta[,1] # primera columna, que són els optims.
rec3 <- oumR3$theta[,1] # primera columna, que són els optims.

theta <- rbind(rec1,rec2,rec3)
colnames(theta) <- c("AFC","AFI","TFC","TFI","TOC")
theta <- data.frame(theta)
theta
```

```{r}
library("ggplot2")
library("dplyr")
library("hrbrthemes")
library("viridis")
library("tidyverse")
library("reshape2")
```

```{r}
melt(theta) 

theta <- melt(theta) 

colnames(theta) <- c("ecoAll","optims")
```

```{r}
# Plot
  ggplot(theta, aes(x=ecoAll, y=optims, fill=ecoAll)) +
    geom_violin(width=1.4) +
    geom_boxplot(width=0.1, color="grey", alpha=0.2) +
    scale_fill_viridis(discrete = TRUE) +
    theme_ipsum()  +
    labs(x = NULL,y = "Phenotypic optima (relative brain size)") +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
     theme(axis.title.x = NULL,
        axis.text.x = NULL,
        axis.title.y = element_text(size = 14))+
  scale_fill_manual(values=c("red","olivedrab3","darkgreen","blue","purple"))
  
```


